---
title: "zhang_garnett"
output: html_document
date: "2024-03-21"
---

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
library(garnett)
library(Seurat)
library(ggplot2)
library(org.Hs.eg.db)
```

#read in the macrophage subset from the 20 combined BPH samples
```{r}
seurat.mf <- readRDS("macs.integratedADTnorm.RDS")
DefaultAssay(seurat.mf) <- "RNA"
```

# create a new CDS object
```{r}
cds_mf<-as.CellDataSet(seurat.mf)
cds_mf<- estimateSizeFactors(cds_mf)
saveRDS(cds_mf, "cds_mf_v1.RDS")

cds_mf<-readRDS("cds_mf_v1.RDS")
```

#read in marker genes, check and plot them to identify genes not found in the database
```{r}
marker_file<-"zhang_garnett_markeers_orig"
raw_marker_check <- check_markers(cds_mf, marker_file,
                              db=org.Hs.eg.db,
                              cds_gene_id_type = "SYMBOL",
                              marker_file_gene_id_type = "SYMBOL")

plot_markers(raw_marker_check)

#Genes not found in db: CTD-2337-J16.1, RNA28S5
#modified marker file and replotted
#maybe also need to remove ANXA2 bc its a lil ambiguous


#training the clasifier and identifying feature genes
zhang_classifier<-train_cell_classifier(cds_mf, marker_file = marker_file, db=org.Hs.eg.db,
                                         cds_gene_id_type = "SYMBOL",
                                         num_unknown = 50,
                                         marker_file_gene_id_type = "SYMBOL")

feature_genes <- get_feature_genes(zhang_classifier,
                                   node = "root",
                                   db = org.Hs.eg.db)
head(feature_genes)

saveRDS(zhang_classifier,"zhang_classifier.RDS")
```

#using classify cells function on our BPH sample object
```{r}
cds_mf_v2 <- classify_cells(cds_mf, zhang_classifier,
                           db = org.Hs.eg.db,
                           cluster_extend = TRUE,
                           cds_gene_id_type = "SYMBOL")

saveRDS(cds_mf_v2, "zhang_cds_mf_v2.RDS")
```

#inserting le cell column into the seurat object
```{r}
seurat.mf$RA_cell_type<-cds_mf_v2$cell_type
print(table(seurat.mf@meta.data$RA_cell_type))
```
#plotting + visualization
```{r}
cols <- c( "#C77CFF","#FF68A1","#00A9FF",  "#E68613", "gray")

plot(DimPlot(seurat.mf, group.by = "RA_cell_type", cols = cols, label = F) + ggtitle("Rheumatoid Arthritis Macrophages in BPH Subclusters"))

pdf("RA_MFs.pdf")
plot(DimPlot(seurat.mf, group.by = "RA_cell_type", cols = cols, label = F) + ggtitle("Rheumatoid Arthritis Macrophages in BPH Subclusters"))
dev.off()
```

#printing the counts and percentages 
```{r}
ra_mf_counts <-table(seurat.mf$RA_cell_type, seurat.mf$seurat_clusters)
ra_mf_counts
ra_mf_perc <- round((prop.table(ra_mf_counts, margin = 2) * 100), 2)
ra_mf_perc
```





